Data for ms. 1

### 2017-04-28 ms1/ms1.Rmd
ms1

summarize_by = function(usage, by) {
  group_by_(usage, by) %>%
  summarize(n_claims=sum(n_claims), n_bene=n()) %>%
  mutate(cpkp=n_claims*1000/n_bene) %>%
  select_(by, 'cpkp')
}
```

@@

load_data = function(year) {
  usage = read_tsv(sprintf('usage_%i.tsv', year))
  pde = read_tsv(sprintf('pde_%i.tsv', year))

  list(usage=usage, pde=pde)
}
```

```{r single_condition_model}
single_condition_model = function(x, condition) {
  filter(x, age >= 67) %>%
  mutate(age_over_67 = age - 67) %>%
  rename_(.dots=setNames(condition, 'has_condition')) %>%
    glm(n_claims ~ age_over_67 + is_female + has_condition,
        family=quasipoisson(link='identity'), data=.) %>%
    tidy() %>%
    mutate(condition=condition)
}

single_condition_models = function(usage) {
  lapply(conditions, function(cond) single_condition_model(usage, cond)) %>%
    bind_rows()
}
```

```{r analyze_function}
age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 95, 100))
claims_group_f = function(x) bounds_group(x, c(0, 1, 2, 5, 10, 20, Inf))

@@

  total_n_bene = nrow(dat$usage)
  total_n_pde = nrow(dat$pde)

@@

  # from here on, the denominators are taken from the beneficiary data grouped
  # in the same way as the consumption data, so we can use the summarize_by function
  claims_by_sex = summarize_by(dat$usage, 'is_female')

  # then by age group
  # (the input to summarize_by is modified, since we want to group by age *group*,
  # not just by raw age)
  claims_by_age = dat$usage %>%
    group_by(age) %>%
    summarize(n_bene=n(),
              mean_n_claims=mean(n_claims),
              sd_n_claims=sd(n_claims),
              mean_n_claims_0cc=mean(n_claims[n_cc==0]),
              sd_n_claims_0cc=sd(n_claims[n_cc==0]),
              mean_n_cc=mean(n_cc),
              sd_n_cc=sd(n_cc)) %>%
    mutate(sem_n_claims=sd_n_claims/sqrt(n_bene),
           sem_n_claims_0cc=sd_n_claims_0cc/sqrt(n_bene),
           sem_n_cc=sd_n_cc/sqrt(n_bene))

  chars_by_age_group = dat$usage %>%
    mutate(age_group=age_group_f(age)) %>%
    group_by(age_group) %>%
    summarize(n_bene=n(),
              mean_n_claims=mean(n_claims),
              mean_n_cc=mean(n_cc)) %>%
    mutate(cpkp=mean_n_claims*1000) %>%
    select(age_group, n_bene, cpkp, mean_n_cc)

  claims_by_region = summarize_by(dat$usage, 'region')

  chars_by_consumption = dat$usage %>%
    mutate(claims_group=claims_group_f(n_claims)) %>%
    group_by(claims_group) %>%
    summarize(`no. beneficiaries`=n(),
              `mean age`=mean(age),
              `% female`=mean(is_female) * 100,
              `mean no. chronic conditions`=mean(n_cc))

  condition_model = single_condition_models(dat$usage)

  list(totals=totals,
       claims_by_class=claims_by_class,
       top_abx=top_abx,
       claims_by_sex=claims_by_sex,
       claims_by_age=claims_by_age,
       chars_by_age_group=chars_by_age_group,
       claims_by_region=claims_by_region,
       chars_by_consumption=chars_by_consumption,
       condition_model=condition_model) %>%
    lapply(function(x) mutate(x, year=year))
}

@@

@@

@@

```{r fig1_claims_by_age}
claims_by_age %>%
  filter(year==2011) %>%
  ggplot(aes(x=age,
             y=mean_n_claims,
             ymin=mean_n_claims-sem_n_claims,
             ymax=mean_n_claims+sem_n_claims)) +
  geom_point() +
  geom_errorbar() +
  theme_minimal() +
  scale_x_continuous(breaks=c(66, 70, 80, 90, 100))
```

@@

```{r}
f = function(df) {
  stopifnot(nrow(df) == 1)
  data_frame(group=c('all', '0cc'),
             mean_n_claims=c(df$mean_n_claims, df$mean_n_claims_0cc),
             sem_n_claims=c(df$sem_n_claims, df$sem_n_claims_0cc))
}

claims_by_age %>%
  filter(year==2011, between(age, 66, 96)) %>%
  select(age, mean_n_claims, mean_n_claims_0cc) %>%
  gather('group', 'mean_n_claims', mean_n_claims, mean_n_claims_0cc) %>%
  ggplot(aes(x=age,
             y=mean_n_claims,
             color=group,
             group=group)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE) +
  theme_minimal() +
  ylab('claims per person') +
  scale_x_continuous(breaks=c(66, 70, 80, 90, 96)) +
  scale_y_continuous(breaks=c(0, 0.5, 1, 1.5, 2.0), limits=c(0, 2))
```

>>>>

summarize_by = function(bene, by) {
  group_by_(bene, by) %>%
  summarize(n_claims=sum(n_claims), n_bene=n()) %>%
  mutate(cpkp=n_claims*1000/n_bene) %>%
  select_(by, 'cpkp')
}
```

@@

load_data = function(year) {
  bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
    filter(between(age, 66, 96), hmo_months==0) %>%
    mutate(is_female=sex=='female', is_white=race=='white') %>%
    left_join(regions, by='state')

  pde = read_tsv(sprintf('../pde_%i.tsv', year))

  # get total numbers of PDEs
  n_claims = pde %>% count(bene_id) %>% rename(n_claims=n)
  bene %<>%
    left_join(n_claims, by='bene_id') %>%
    replace_na(list(n_claims=0))

  cc = read_feather(sprintf('../cc_%i.feather', year)) %>%
    mutate(n_cc=rowSums(select_(., '-bene_id')))

  bene %<>% left_join(cc, by='bene_id')

  cu = read_tsv('../cu_%i.tsv') %>%
    select(bene_id, n_visits=hospital_op_visits)

  bene %<>% left_join(cu, by='bene_id')

  list(bene=bene, pde=pde)
}
```

```{r single_condition_model}
single_condition_model = function(x, condition) {
  filter(x, age >= 67) %>%
  rename_(.dots=setNames(condition, 'has_condition')) %>%
    glm(n_claims ~ age + is_female + has_condition,
        family=quasipoisson(link='identity'), data=.) %>%
    tidy() %>%
    mutate(condition=condition)
}

single_condition_models = function(bene) {
  lapply(conditions, function(cond) single_condition_model(bene, cond)) %>%
    bind_rows()
}
```

```{r analyze_function}
age_group_f = function(x) bounds_group(x, c(65, 70, 75, 80, 85, 90, 96))
claims_group_f = function(x) bounds_group(x, c(0, 1, 2, 5, 10, 20, Inf))

@@

  total_n_bene = nrow(dat$bene)
  total_n_pde = nrow(dat$pde)

@@

  # from here on, the denominators are taken from the beneficiary data grouped
  # in the same way as the consumption data, so we can use the summarize_by function
  claims_by_sex = summarize_by(dat$bene, 'is_female')

  # then by age group
  # (the input to summarize_by is modified, since we want to group by age *group*,
  # not just by raw age)
  claims_by_age = dat$bene %>%
    bind_rows(. %>% mutate(group='all'),
              . %>% filter(n_cc==0) %>% mutate(group='no_cc'),
              . %>% filter(n_visits==0) %>% mutate(group='no_visits')) %>% 
    group_by(group, age) %>%
    summarize(n_bene=n(),
              mean_n_claims=mean(n_claims),
              sd_n_claims=sd(n_claims),
              mean_n_cc=mean(n_cc),
              sd_n_cc=sd(n_cc),
              mean_n_visits=mean(n_visits),
              sd_n_visits=sd(n_visits)) %>%
    mutate(sem_n_claims=sd_n_claims/sqrt(n_bene),
           sem_n_cc=sd_n_cc/sqrt(n_bene),
           sem_n_visits=sd_n_visits/sqrt(n_bene))

  chars_by_age_group = dat$bene %>%
    mutate(age_group=age_group_f(age)) %>%
    group_by(age_group) %>%
    summarize(n_bene=n(),
              mean_n_claims=mean(n_claims),
              mean_n_cc=mean(n_cc)) %>%
    mutate(cpkp=mean_n_claims*1000) %>%
    select(age_group, n_bene, cpkp, mean_n_cc)

  claims_by_region = summarize_by(dat$bene, 'region')

  claims_by_race = dat$bene %>%
    mutate(race=if_else(race %in% c('white', 'black', 'Hispanic'), race, 'other')) %>%
    summarize_by('race')

  chars_by_consumption = dat$bene %>%
    mutate(claims_group=claims_group_f(n_claims)) %>%
    group_by(claims_group) %>%
    summarize(`no. beneficiaries`=n(),
              `mean age`=mean(age),
              `% female`=mean(is_female) * 100,
              `mean no. chronic conditions`=mean(n_cc))

  condition_model = single_condition_models(dat$bene)

  list(totals=totals,
       claims_by_class=claims_by_class,
       top_abx=top_abx,
       claims_by_sex=claims_by_sex,
       claims_by_race=claims_by_race,
       claims_by_age=claims_by_age,
       chars_by_age_group=chars_by_age_group,
       claims_by_region=claims_by_region,
       chars_by_consumption=chars_by_consumption,
       condition_model=condition_model) %>%
    lapply(function(x) mutate(x, year=year))
}

@@

claims_by_race = data_by_field('claims_by_race') %>%
  mutate(cpkp=round(cpkp)) %>%
  spread(year, cpkp) %T>%
  write_tsv('tables/tmp_race.tsv') %>%
  kable

@@

## Consumption by race

`r claims_by_race`

@@

```{r fig1_claims_by_age}
claims_by_age %>%
  filter(year==2011) %>%
  ggplot(aes(x=age,
             y=mean_n_claims,
             ymin=mean_n_claims-sem_n_claims,
             ymax=mean_n_claims+sem_n_claims,
             color=group)) +
  geom_point() +
  geom_linerange() +
  theme_minimal() +
  scale_x_continuous(breaks=c(66, 70, 80, 90, 96))
```

@@

```{r stratified_claims_by_age}
claims_by_age %>%
  filter(year==2011, between(age, 66, 96)) %>%
  select(age, mean_n_claims, mean_n_claims_0cc) %>%
  gather('group', 'mean_n_claims', mean_n_claims, mean_n_claims_0cc) %>%
  ggplot(aes(x=age,
             y=mean_n_claims,
             color=group,
             group=group)) +
  geom_point(shape=1, cex=5) +
  geom_smooth(method='lm', se=FALSE) +
  theme_minimal(base_size=24) +
  ylab('claims per person') +
  scale_x_continuous(breaks=c(66, 70, 80, 90, 96)) +
  scale_y_continuous(breaks=c(0, 0.5, 1, 1.5, 2.0), limits=c(0, 2))

claims_by_age %>%
  filter(year==2011, between(age, 67, 96)) %>%
  ggplot(aes(x=age, y=mean_n_cc)) +
  geom_point(shape=1, size=6) +
  geom_smooth(method='lm', se=FALSE, col='black') +
  theme_minimal(base_size=24) +
  scale_x_continuous(breaks=c(67, 70, 80, 90, 96)) +
  ylab('mean no. chronic conditions') +
  ylim(0, 6)
```
